%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \miguel{Indeed,having a section in the final document where we give definitions of "quality", "software", "research software", etc. is
% quite convenient. It doesn't mean other definitions are not possible, but we need to settle clearly what we're discussing in our own context.}
% Criteria of quality (brainstorming, to complete by everybody)

% \miguel{For scientific research: reviewed by other peers, reproducible (data + sources, matching the article), Open Access.}

% \dg{We should look at the FAIR4RS reports for the alignment for quality dimensions. In a sense, this is what FAIR aims (at a basic level).}

% \miguel{One quality criteria: reproducibility. Define "reproducibility" and "repeatability". Different levels of reproducibility. Bottom level:
% black box (being able to obtain the same results given the same inputs). Higher level: full access to a well-commented source code, and checking
% that the pseudocode matches well the implementation. In the case of a scientific publication, checking that the implementation matches exactly
% what the descriptions in the article.}

% \dg{We should first define what we understand by software. We can reuse the definition in FAIR4RS.}

% These refer to the work done on identifying Quality Models and Attributes - table under discussion as of March 2023
% DONE -> \dg{Missing the initial table here, although we had only summaries. We should see how each of those papers defined software quality dimensions.}

% DONE -> \eli{Concerning Quality Definition, there are various software models and standards that can be referred.
% However, the most recent ones are ISO and IEEE. Existing models and standards provide a full set of software characteristics: unfortunately
% they quite often have different meaning and do not have a match with software metrics used to measure software characteristics.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\subsection{Software Quality characteristics}

\begin{center}
    \tablefirsthead{
        \hline
        \textbf{Characteristic} & \textbf{Definition} & \textbf{Notes} \\
        \hline
    }
    \tablehead{
        \hline
        \textbf{Characteristic} & \textbf{Definition} & \textbf{Notes} \\
        \hline
    }
    \tabletail{\hline}
    \tablelasttail{\hline}
    
    \bottomcaption{Software Quality characteristics are taken from ISO/IEC 25010:2011(E)~\cite{iso_25010_2011_2017} except Scalability and Supportability.
    Microsoft: \url{https://docs.microsoft.com/en-us/previous-versions/msp-n-p/ee658094(v=pandp.10)}.}
    \label{tab:sq_char}
    
    \small
    \begin{supertabular}{|p{0.15\linewidth}|p{0.4\linewidth}|p{0.4\linewidth}|}
    
    \hline
    
    Functional suitability &
    Degree to which a product or system provides functions that meet stated and implied needs when used under specified conditions. &
    -
    \\ \hline
    
    Availability &
    Degree to which a system, product or component is operational and accessible when required for use. &
    Adapted from ISO/IEC/IEEE 24765
    \\ \hline
    
    Reliability &
    Degree to which a system, product or component performs specific functions under specified conditions for a specified period of time. &
    Adapted from ISO/IEC/IEEE 24765. Limitations in reliability are due to faults in requirements, design and implementation, or due to contextual changes.
    \\ \hline
    
    Time behaviour &
    Degree to which the response and processing times and throughput rates of a product or system, when performing its functions, meet requirements. &
    -
    \\ \hline
    
    Performance &
    Performance relative to the amount of resources used under stated conditions. &
    Resources can include other software products, the software and hardware configuration of the system, and materials (e.g. print paper, storage media).
    \\ \hline
    
    Ease of use (Usability) &
    Degree to which a product or system can be used by specified users to achieve specific goals with effectiveness, efficiency and satisfaction in a specified context of use. &
    Adapted from ISO 9241-210. Usability can either be specified or measured as a product quality characteristic in terms of its sub-characteristics, or specified or measured directly by measures that are a subset of quality in use.
    \\ \hline
    
    Fault tolerance &
    Degree to which a system, product or component operates as intended despite the presence of hardware or software faults. &
    Adapted from ISO/IEC/IEEE 24765.
    \\ \hline
    
    Security &
    Degree to which a product or system protects information and data so that persons or other products or systems have the degree of data access appropriate to their types and levels of authorisation. &
    As well as data stored in or by a product or system, security also applies to data in transmission. Survivability (the degree to which a product or system continues to fulfil its mission by providing essential services in a timely manner in spite of the presence of attacks) is covered by recoverability (4.2.5.4). Immunity (the degree to which a product or system is resistant to attack) is covered by integrity (4.2.6.2). Security contributes to trust (4.1.3.2).
    \\ \hline
    
    Maintainability &
    Degree of effectiveness and efficiency with which a product or system can be modified by the intended maintainers. &
    Modifications can include corrections, improvements or adaptation of the software to changes in environment, and in requirements and functional specifications. Modifications include those carried out by specialised support staff, and those carried out by business or operational staff, or end users. Maintainability includes installation of updates and upgrades. Maintainability can be interpreted as either an inherent capability of the product or system to facilitate maintenance activities, or the quality in use experienced by the maintainers for the goal of maintaining the product or system.
    \\\hline
    
    Recoverability &
    Degree to which, in the event of an interruption or a failure, a product or system can recover the data directly affected and re-establish the desired state of the system. &
    Following a failure, a computer system will sometimes be down for a period of time, the length of which is determined by its recoverability.
    \\ \hline
    
    Operability / Manageability &
    Degree to which a product or system has attributes that make it easy to operate and control. &
    Operability corresponds to controllability, (operator) error tolerance and conformity with user expectations as defined in ISO 9241-110.
    \\ \hline
    
    Resource utilisation &
    Degree to which the amounts and types of resources used by a product or system, when performing its functions, meet requirements. &
    Human resources are included as part of efficiency (4.1.2).
    \\ \hline
    
    Safety &
    Degree to which a product or system mitigates the potential risk to people in the intended contexts of use. &
    -
    \\ \hline
    
    Interoperability &
    Degree to which two or more systems, products or components can exchange information and use the information that has been exchanged. &
    Based on ISO/IEC/IEEE 24765.
    \\ \hline
    
    Attractiveness &
    Renamed as user interface aesthetics. Degree to which a user interface enables pleasing and satisfying interaction for the user. &
    This refers to properties of the product or system that increase the pleasure and satisfaction of the user, such as the use of colour and the nature of the graphical design.
    \\ \hline
    
    Compatibility &
    Degree to which a product, system or component can exchange information with other products, systems or components, and/or perform its required functions, while sharing the same hardware or software environment. &
    Adapted from ISO/IEC/IEEE 24765.
    \\ \hline
    
    Installability &
    Degree of effectiveness and efficiency with which a product or system can be successfully installed and/or uninstalled in a specified environment. &
    -
    \\ \hline
    
    Technical accessibility &
    Degree to which a product or system can be used by people with the widest range of characteristics and capabilities to achieve a specified goal in a specified context of use. &
    The range of capabilities includes disabilities associated with age. Accessibility for people with disabilities can be specified or measured either as the extent to which a product or system can be used by users with specified disabilities to achieve specified goals with effectiveness, efficiency, freedom from risk and satisfaction in a specified context of use, or by the presence of product properties that support accessibility.
    \\ \hline
    
    Portability / Adaptability &
    Degree of effectiveness and efficiency with which a system, product or component can be transferred from one hardware, software or other operational or usage environment to another. &
    Adapted from ISO/IEC/IEEE 24765. Portability can be interpreted as either an inherent capability of the product or system to facilitate porting activities, or the quality in use experienced for the goal of porting the product or system.
    \\ \hline
    
    Modifiability &
    Degree to which a product or system can be effectively and efficiently modified without introducing defects or degrading existing product quality. &
    Implementation includes coding, designing, documenting and verifying changes. Modularity (4.2.7.1) and analysability (4.2.7.3) can influence modifiability. Modifiability is a combination of changeability and stability.
    \\ \hline
    
    Reusability &
    Degree to which an asset can be used in more than one system, or in building other assets. &
    -
    \\ \hline
    
    Scalability &
    1.) Scalability is the ability of a system to either handle increases in load without impact on the performance of the system, or the ability to be readily enlarged $OR$ 2.) Scalability is the capability of algorithms, protocols, and applications to efficiently handle a growing amount of work or the demand of increasing its performance (according to some metrics) by adding resources to the system in which the software is running. Resources can be added to the single nodes (vertical scalability) and to the system as a whole (horizontal scalability). &
    1.) Microsoft, 2.) Massimo Torquati
    \\ \hline
    
    Supportability &
    1.) Supportability is the ability of the system to provide information helpful for identifying and resolving issues when it fails to work correctly $OR$ 2.) Existence of a helpdesk or issue tracking, bug reporting, enhancements and general support.  &
    1.) Microsoft, 2.) Mario David
    \\ \hline
    
    Testability &
    Degree of effectiveness and efficiency with which test criteria can be established for a system, product or component and tests can be performed to determine whether those criteria have been met. &
    Adapted from ISO/IEC/IEEE 24765.
    \\ \hline
    
    Confidentiality &
    Degree to which a product or system ensures that data are accessible only to those authorised to have access. &
    -
    \\ \hline
    
    \end{supertabular}
\end{center}
    
\subsection{Software Quality attributes}

The metrics and attributes where subdivided into four categories and for each metric or attribute,
a new codename was created by the EOSC Task Force:

\begin{itemize}
    \item Source Code Metrics (\textbf{EOSC-SCMet}):
    \item Time metrics (\textbf{EOSC-TMet}):
    \item Qualitative Attributes (\textbf{EOSC-Qual}):
    \item DevOps - SW release and management Attributes (\textbf{EOSC-SWRelMan}):
    \item DevOps - Testing Attributes (\textbf{EOSC-SWTest}):
    \item Service Operability Attributes (\textbf{EOSC-SrvOps}):
\end{itemize}

In the Appendix is show in bold the EOSC Task Force codename, the name and
one or more characteristics defined in the section "Software Quality characteristics", it is
followed by a table containing the source information about the metric or attribute, i.e. the
reference paper or article, the codename defined there and its definition. In several cases,
the same (or similar) metric or attribute appears in different papers.
