\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[a4paper, total={16.5cm, 22.3cm}]{geometry}
\usepackage{url}
\usepackage{verbatim}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amssymb,amsfonts,textcomp}
\usepackage{color}
\usepackage{array}
\usepackage{hhline}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{supertabular}
\usepackage[table,xcdraw,usenames,dvipsnames]{xcolor}
\usepackage{parskip}
\usepackage[backend=biber,style=numeric]{biblatex}
\usepackage{csquotes}
\usepackage{orcidlink}
\usepackage[toc]{appendix}
\addbibresource{EOSC_IQRS.bib}

\title{Task Force Sub Group 3 - Review of Software Quality Attributes and Characteristics}
\author{
    David, Mario (LIP) \orcidlink{0000-0003-1802-5356} \and
    Colom, Miguel (ENS Paris-Saclay) \orcidlink{0000-0003-2636-0656} \and
    Garijo, Daniel (UPM) \orcidlink{0000-0003-0454-7145} \and
    Castro, Leyla Jael (ZB MED) \orcidlink{0000-0003-3986-0510} \and
    Louvet, Violaine (CNRS) \orcidlink{0000-0002-8742-8952} \and
    Ronchieri, Elisabetta (INFN/CNAF) \orcidlink{0000-0002-7341-6491} \and
    Torquati, Massimo (UNIPI) \orcidlink{0000-0001-6323-3459} \and
    Cano, Laura del (CSIC/CNB) \orcidlink{0000-0003-0981-2040} \and
    Leong, Cerlane (CSCS)  \orcidlink{0000-0001-8241-6277} \and
    Van den Bossche, Maxime (KU Leuven) \orcidlink{0000-0002-0938-0156} \and
    Campos, Isabel (CSIC/IFCA) \orcidlink{0000-0002-9350-0383}
}

\date{\today}
\begin{document}
\maketitle
\tableofcontents

\newpage
\section{Quality Characteristics}

Software Quality characteristics are taken from ISO/IEC 25010:2011(E)~\cite{iso_25010_2011_2017} except Scalability and
Supportability \cite{microsoft_2010}.

\input{qual-char}

\newpage
\section{Quality Attributes}

We follow the methodology for the survey, proposed by Kitchenham and Charters \cite{keele2007guidelines} which has the following steps:

\begin{enumerate}
    \item Source selection and search: We have searched in the Scopus dataset, including the top five journals in
    software engineering related to software \footnote{\url{https://research.com/journals-rankings/computer-science/software-programming}}
    and articles of the  "International Conference on Software Engineering", one of the top venues for Software engineering. We also added documentclass
    and web resources that the Task Force subgroup considered relevants.
    The search included the keywords "software quality" in the title of the target publications.
    \item Inclusion and exclusion criteria: Excluded journals not in the SE domain. Excluded articles not written in English.
    \item Selection procedure: Skim article titles and abstracts. The process was performed by 2-3 people. Final list was agreed upon by the group through discussion
    about the relevance of the paper and analysis if that paper contains or proposes Software Quality attributes.
    \item Review process: After following the selection procedure, we ended up with 19 articles, which have been reviewed in this survey. Some of the articles
    refer to the ISO/IEC 25010:2011(E)~\cite{iso_25010_2011_2017} or to it's precursor ISO/IEC 9126, have been grouped together.
\end{enumerate}

Journals: IEEE Transactions on Software Engineering, Empirical Software Engineering, Journal of Systems and Software, Software \& Systems Modeling, 
Information and Software Technology, IEEE Software, Software Quality Journal. Query used:

\tiny
\begin{verbatim}
    TITLE ( software  AND quality )  AND  
        ( LIMIT-TO ( EXACTSRCTITLE ,  "Software Quality Journal" )  
        OR  LIMIT-TO ( EXACTSRCTITLE ,  "Proceedings International Conference On Software Engineering" ) 
        OR  LIMIT-TO ( EXACTSRCTITLE ,  "IEEE Transactions on Software Engineering" )
        OR  LIMIT-TO ( EXACTSRCTITLE ,  "Empirical Software Engineering" ) 
        OR  LIMIT-TO ( EXACTSRCTITLE ,  "Journal of Systems and Software" ) 
        OR  LIMIT-TO ( EXACTSRCTITLE ,  "Software & Systems Modeling" ) 
        OR  LIMIT-TO ( EXACTSRCTITLE ,  "Information and Software Technology" )  
        OR  LIMIT-TO ( EXACTSRCTITLE ,  "IEEE Software" )   
        )  AND  ( LIMIT-TO ( SUBJAREA ,  "COMP" )  OR  LIMIT-TO ( SUBJAREA ,  "ENGI" ) )  
\end{verbatim}
\small

As a result, we obtained 272 results. Additional filtering: excluding papers with no abstracts, proceedings/workshop summary,
and removed those which did not seem related by browsing the abstract and title. Also, removed those papers that did not seem to propose
quality dimensions (e.g., if they talk about practices).

The metrics and attributes where subdivided into four categories and for each metric or attribute,
a new codename was created by the EOSC Task Force:

\begin{itemize}
    \item Source Code Metrics (\textbf{EOSC-SCMet}):
    \item Time metrics (\textbf{EOSC-TMet}):
    \item Qualitative Attributes (\textbf{EOSC-Qual}):
    \item DevOps - SW release and management Attributes (\textbf{EOSC-SWRelMan}):
    \item DevOps - Testing Attributes (\textbf{EOSC-SWTest}):
    \item Service Operability Attributes (\textbf{EOSC-SrvOps}):
\end{itemize}

Research Software levels correspond to the four user stories defined in subgroup 1 (Research Software Lifecycle) of
this Task Force (Infrastructure for Quality Research Software) \footnote{\url{https://docs.google.com/document/d/13TzhxNpGLtbFWYOmWSKkEHFC22ICmVpB/}},
and are the following:

\begin{enumerate}
    \item \textbf{Individual} creating research software for own use (e.g. a PhD student): easy to implement, good practice for research software at any level.
    \item \textbf{Team} creating an application or workflow for use within the team: easy to implement, good practice for research software at any level, useful
    for some basic coordination when more than one person participates.
    \item \textbf{OSS} A team / community developing (possibly broadly applicable) open source research software or service platform:
    Open Source Software in general, all other cases.
\end{enumerate}

\input{09-appendix}

\newpage
\printbibliography

\end{document}
